---
title: "Defining and applying data quality rules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Defining and applying data quality rules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(knitr)
```

As part of the 'auto.info.report' package we have supplied a framework for building and applying your own data quality rules.  In the package we consider two distinct types of data quality measure:

1. Those that apply to each value of a variable, e.g. are observations missing, or aren't within the expected domain.
2.  Those that apply at the variable level, e.g. are the variable highly skewed, or have > 10% missing values.

The core of this behavior are the `new_rule` and `new_variable_rule` functions - which are used for defining the *expected* behavior i.e. `TRUE` implies quality.  Each function takes three parameters:

* A human readable name
* A function to define the prefered behaviour
* [optional] a `tidyselect` function of variables to apply the rules to.  By default rules will apply  to every variable.

```{r setup}
library(auto.info.report)
library(tidyselect)

rule.1 <- new_rule("Non-missing values", function(x) !is.na(x))

variable.rule.1 <- new_variable_rule("Low skew", 
                  function(x) abs(e1071::skewness(x)) < 2, 
                  where(is.numeric))

rule.list <- list(rule.1, variable.rule.1)
```

With rules defined, we call `apply_rules` and `dq_rule_report` to produce a list with two items:

* `ValueRules` to summarize the rules defined via `new_rule`
* `VariableRules` to summarize the rules defined via `new_variable_rule`

``` {r, results="asis"}
dq.list <- apply_rules(iris, rule.list)
rule.report <- dq_rule_report(dq.list)
```

The `ValueRules` summarize what percentage of each variable follows the rule:

```{r}
kable(rule.report$ValueRules)
```

The `VaraiableRules` summarize if the variable(s) follows the rule:

``` {r}
kable(rule.report$VariableRules)
```

### Useful utility functions

To aid in construction of DQ rules, we supply a few helper functions:

* `select_numerics` for rules to apply to numeric variables
* `select_booleans` for rules to apply to logical variables
* `select_dates`  for rules to apply to date variables
* `select_POSIXct` for rules to apply to datetime variables
* `select_character` for rules to apply to string variables
* `select_factor` for rules to apply to factor variables
* `select_categorical` for rules to apply to string and factor variables

while the general family of `tidyselect` verbs can be used, e.g.:

* `starts_with`
* `ends_with`
* `matches`
* `contains`

To aid in creating the rules we supply a similar set of functions:

* `greater_than.f` and `greater_than_zero`
* `less_than.f` and `less_than_zero`
* `not.na`
* `one_of.f`
* `limited_cardinality`
* `within_interval`
