---
title: "Intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(MI.report)
```

When working with multiple data sets it can be helpful to have a standardized 
data pipeline for reporting summary metrics and making the reports openly 
available.  This package serves as an API for producing data-type dependent 
summaries of all variables in a data set rendered as html reports for access 
via web hosting e.g. github pages.

The key entry to the package is via the `new_data_report` function: 

```{r setup, eval=F}
library(MI.report)
dr <- new_data_report(path = "~/Data.Report")
```

which takes one argument - the system path to the desired directory for the 
data report files.  The `new_data_report` function return a `DataReport` class 
which handles the construction of vignettes and pkgdown features.

In order to write a vignette using the supplied template - `DataReport` expects
a list of data frames, one for each data types defined (currently `numeric`, 
`date`,  `datetime` and `other`) where each data type have their own summary 
metrics.  The function `typed_summary_metrics` has been supplied as an easy 
binding to the default summary metrics, and below we demonstrate it being 
applied to the standard `iris` data set.

``` {r}
`Iris Data` <- list(
  typed.report = typed_summary_metrics(iris)
)
``` 

To template a new vignette off of this summary - call the `add_vigentte` method
of the earlier generated `DataReport`, passing the metrics as the data argument.
The package will then store this data in the desired data report project, and 
create a vignette based on the values.

``` {r, eval=F}
dr$add_vignette(name="Iris Data", 
                data=`Iris Data`, description="Example data set - IRIS ")
```

To launch the website and check the new vignette  evaluate the `build_site`
method.

``` {r, eval=F}
dr$build_site()
```
